### 2015-05-05 background.tex
more updates in the lockset analysis part of the background section

The above example also showcases the main limitation of happens-before. If the technique explores \emph{only} this specific thread interleaving, it will miss a potential data race between statements $B$ and $G$, because $B \rightarrow G$ in this specific schedule. Thus, a tool based on happens-before requires a scheduler that produces as many different thread interleavings as possible to increase execution path coverage~\cite{savage1997eraser}. Achieving full coverage with tools based on happens-before is infeasible though, because of the exponential number of possible thread interleavings in realistic concurrent programs. Furthermore, attempting to explore a large number of thread schedules quickly leads to scalability issues~\cite{musuvathi2008finding}, a problem known as state space explosion. Thus, previous work based on happens-before typically explore paths under a certain bound, such as the number of allowed context switches~\cite{qadeer2004kiss}.

@@

\ADComment{What is missing from the explanation, and from the example you give, is the idea that with a dynamic approach, you run the program and a thread operates in a truly concurrent environment.  You thus only get to see what the thread did on the particular schedule that was followed.  I also think you should use a two-threaded example, instead of a one-threaded example, to show that although thread $T$ might consistently use mutex $M$ to protect variable $A$, this is no good if thread $U$ accesses variable $A$ without holding $M$.}

@@

\ADComment{I folded some of the discussion of the example into the above more general description of locksets.  Assuming you do produce a two-threaded example, please put in a brief explanation of it.}

In contrast to analyses based on happens-before, lockset analysis has the potential to scale well, because it does not need to reason about all possible thread interleavings to achieve good code coverage. The technique, though, suffers from imprecision (i.e. can report false bugs), because a violation of the locking discipline does not always correspond to a real data race~\cite{savage1997eraser, pozniansky2003efficient, o2003hybrid, elmas2007goldilocks, flanagan2009fasttrack}. Furthermore, code coverage in dynamic lockset analysis tools, such as Eraser, is limited by the number of explored execution paths. To counter these limitations, some techniques~\cite{pratikakis2006locksmith, voung2007relay} have explored the idea of applying lockset analysis in a static context, using dataflow analysis. In this paper, we also present a static lockset analysis method that involves generating verification conditions to be discharged to a theorem prover.

>>>>

The above example also showcases the main limitation of happens-before. If the technique explores \emph{only} this specific thread interleaving, it will miss a potential data race between statements $B$ and $G$, because $B \rightarrow G$ in this specific schedule. Thus, a tool based on happens-before requires a scheduler that produces as many different thread interleavings as possible to increase execution path coverage~\cite{savage1997eraser}. Achieving full coverage with tools based on happens-before is infeasible though, because of the exponential number of possible thread interleavings in realistic concurrent programs. Moreover, attempting to explore a large number of thread schedules quickly leads to scalability issues~\cite{musuvathi2008finding}, a problem known as state space explosion. Thus, previous work based on happens-before typically explore paths under a certain bound, such as the number of allowed context switches~\cite{qadeer2004kiss}.

@@

@@

\ADComment{I also think you should use a two-threaded example, instead of a one-threaded example, to show that although thread $T$ might consistently use mutex $M$ to protect variable $A$, this is no good if thread $U$ accesses variable $A$ without holding $M$.}

In contrast to analyses based on happens-before, lockset analysis has the potential to scale well, because it does not need to reason about all possible thread interleavings to achieve good code coverage. The technique, though, suffers from imprecision (i.e. can report false bugs), because a violation of the locking discipline does not always correspond to a real data race~\cite{savage1997eraser, pozniansky2003efficient, o2003hybrid, elmas2007goldilocks, flanagan2009fasttrack}. Furthermore, the code coverage in dynamic lockset analyzers, such as Eraser, is limited by the execution paths that are explored under a given scheduler. This is because the tool runs the program in a truly concurrent environment, thus it only gets to see what the thread did on the particular schedule that was followed.

To counter the above limitations, techniques such as Locksmith~\cite{pratikakis2006locksmith} and RELAY~\cite{voung2007relay} have explored the idea of applying lockset analysis in a static context, using dataflow analysis. In this paper, we present a novel static lockset analysis method that involves generating verification conditions, which are then discharged to a theorem prover.
